|

### 2021-09-06 submission/manuscript.rmd
update title

---
title: |
 A model including an interaction of treatment with baseline risk was the optimal
 risk-based approach to predicting absolute treatment benefit
abstract: |
 **Objective:** To compare different risk-based methods predicting individualized
 treatment effects in RCT simulations. **Study Design and Setting:** We simulated
 data using diverse assumptions for a baseline prognostic index of risk (PI) and
 the shape of its intereraction with treatment (none, linear or quadratic). In 
 each sample we predicted absolute benefit using: models with the PI and a 
 constant relative treatment effect; models including an interaction of treatment 
 with the PI; stratification in quarters of the PI; nonlinear transformations of 
 the PI (restricted cubic splines with 3, 4 and 5 knots); an adaptive approach 
 using Akaike’s Information Criterion. We evaluated predictive performance using 
 root mean squared error and measures of discrimination and calibration for 
 benefit. Starting from a base case scenario (sample size 4,250, treatment odds 
 ratio 0.8, AUC of the PI 0.75), we varied the sample size, the treatment effect 
 strength, and the PI's discriminative ability. **Results:** Models including a 
 PI by treatment interaction performed better under most simulation settings. 
 Flexible models required larger sample sizes and higher AUC of the PI to 
 outperform the linear interaction model. The adaptive approach performed 
 similarly to the best-performing method in most scenarios. **Conclusion:** 
 Under most circumstances, a model with a linear interaction of the PI with 
 treatment is the optimal risk-based approach to predicting absolute treatment
 benefit.
authors:
  - name: Alexandros Rekkas
    department: Department of Medical Informatics
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
  - name: Peter R. Rijnbeek
    department: Department of Medical Informatics
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
  - name: Ewout W. Steyerberg
    department: Department of Biomedical Data Sciences
    affiliation: Leiden University Medical Center
    location: Leiden, The Netherlands
  - name: David van Klaveren
    department: Department of Public Health
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
keywords:
  - treatment effect heterogeneity
  - absolute benefit
  - prediction models
output: 
  rticles::arxiv_article:
    keep_tex: true
  bookdown::word_document2:
    reference_docx: reference.docx
geometry: margin=1.0in
date: false
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \usepackage{amsmath}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \newcommand\given[1][]{\:#1\vert\:}
  - \date{}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

>>>>

---
title: |
 Linear interaction of treatment with baseline risk was sufficient for predicting
 individualized benefit
abstract: |
 **Objective:** To compare different risk-based methods predicting individualized
 treatment effects in RCT simulations. **Study Design and Setting:** We simulated
 data using diverse assumptions for a baseline prognostic index of risk (PI) and
 the shape of its intereraction with treatment (none, linear or quadratic). In 
 each sample we predicted absolute benefit using: models with the PI and a 
 constant relative treatment effect; models including an interaction of treatment 
 with the PI; stratification in quarters of the PI; nonlinear transformations of 
 the PI (restricted cubic splines with 3, 4 and 5 knots); an adaptive approach 
 using Akaike’s Information Criterion. We evaluated predictive performance using 
 root mean squared error and measures of discrimination and calibration for 
 benefit. Starting from a base case scenario (sample size 4,250, treatment odds 
 ratio 0.8, AUC of the PI 0.75), we varied the sample size, the treatment effect 
 strength, and the PI's discriminative ability. **Results:** Models including a 
 PI by treatment interaction performed better under most simulation settings. 
 Flexible models required larger sample sizes and higher AUC of the PI to 
 outperform the linear interaction model. The adaptive approach performed 
 similarly to the best-performing method in most scenarios. **Conclusion:** 
 Under most circumstances, a model with a linear interaction of the PI with 
 treatment is the optimal risk-based approach to predicting absolute treatment
 benefit.
authors:
  - name: Alexandros Rekkas
    department: Department of Medical Informatics
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
  - name: Peter R. Rijnbeek
    department: Department of Medical Informatics
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
  - name: Ewout W. Steyerberg
    department: Department of Biomedical Data Sciences
    affiliation: Leiden University Medical Center
    location: Leiden, The Netherlands
  - name: David van Klaveren
    department: Department of Public Health
    affiliation: Erasmus Medical Center
    location: Rotterdam, The Netherlands
keywords:
  - treatment effect heterogeneity
  - absolute benefit
  - prediction models
output: 
  rticles::arxiv_article:
    keep_tex: true
  bookdown::word_document2:
    reference_docx: reference.docx
geometry: margin=1.0in
date: false
toc: false
font-size: 11pt
header-includes:
  - \renewcommand*\familydefault{\sfdefault}
  - \usepackage{setspace}
  - \usepackage{amsmath}
  - \doublespacing
  - \usepackage[left]{lineno}
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \newcommand\given[1][]{\:#1\vert\:}
  - \date{}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---
